Studies of quantum dots

### 2016-09-09 Manuscript/paper.tex
Update spEnergiesQD paper

\documentclass[amsmath, amssymb, aps, floatfix, nofootinbib, preprintnumbers,showpacs, superscriptaddress, twocolumn]{revtex4-1}
\usepackage{bm}
\usepackage[cmyk]{xcolor}
\usepackage{dcolumn}
\usepackage{epsfig}
\usepackage{float}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{subfigure}
\newcommand{\lb}{\left(}
\newcommand{\rb}{\right)}
\newcommand{\ad}{a^{\dagger}}
\newcommand{\Hd}{\hat{H}^{\rm d}}
\newcommand{\Ho}{\hat{H}^{\rm od}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\T}{\hat{T}_{\rm rel}}
\newcommand{\da}{^\dagger}
\newcommand{\lla}{\left\langle}
\newcommand{\rra}{\right\rangle}
\newcommand{\llb}{\lbrace}
\newcommand{\rrb}{\rbrace}
\newcommand{\rv}{\mathbf{r}}

%Jorgen commands
\newcommand{\OP}[1]{{\hat{#1}}}
\newcommand{\ket}[1]{\left| #1 \right>}
\newcommand{\bra}[1]{\left< #1 \right|}
\newcommand{\braket}[2]{\left\langle #1 | #2\right\rangle}

@@

\begin{abstract}
  We present and compare several many-body methods used to study a quantum dot
  system.  We calculate the approximate ground state energy of electrons in a
  circularly symmetric potential well using a harmonic oscillator basis
  optimized via Hartree-Fock.  We further improve the ground state energy
  using in-medium similarity renormalization group (IM-SRG).  Additionally, we
  calculate the addition and removal energies using quasidegenerate
  perturbation theory (QDPT).  Our results are benchmarked against quasi-exact
  diffusion Monte Carlo and Coupled Cluster calculations.
\end{abstract}

@@

Understanding the behaviour of strongly confined electrons is of fundamental
interest for solving many-body problems.  Quantum dots, e.g. electrons
confined in semiconducting heterostructures, are of particular interest since
they exhibit, due to their small size, discrete quantum levels. Under these
conditions, typical quantum phenomena like tunnelling, entanglement and
magnetization can all be observed \cite{reimann2002,engel1993}.  Since quantum
dots are manufactured and designed artificially at the laboratory, their
quantum levels can be tuned to one's needs by changing for instance the
external field, or the size and shape of the system. As a consequence, quantum
dots provide a high level of control for the dynamics and correlation of the
electrons, which makes them perfectly suited to study quantum effects in
practice.  Since their ground state shows similar shell structures and magic
numbers as seen for atoms and nuclei \cite{tarucha1996}, these systems give
the opportunity to study electronic systems without the presence of a nucleus
affecting the electrons.  Apart from their relevance for theoretical research
in quantum physics, quantum dots offer a wide variety of applications: In
particular, their electrical and optical properties make them attractive for
the use in laser technology \cite{strauf2010,5075760} and solar cells
\cite{jenks:013111,doi:10.1021/cr900289f}, but they are also used in quantum
computers \cite{PhysRevA.57.120} and medical imaging \cite{Ben-Ari02042003}.

In order to properly understand the properties of quantum dots and make
theoretical predictions to their behaviour in various applications, it is
necessary to study features like ground state energy and correlation
effects. Since apart from quantum dots consisting of only two electrons or
with specific values of the external field, no analytical solutions exist
\cite{PhysRevA.48.3561}, the development of appropriate few- and many-body
methods is required.

Several \textit{ab initio} methods have been applied to these systems, in
particular variational and diffusion Monte Carlo \cite{PhysRevB.68.035304,%
  PhysRevB.62.8120,PhysRevB.84.115302,PhysRevB.54.4780}, large-scale
diagonalization \cite{JJAP.36.3924,PhysRevB.56.6428,Kvaalcode,rontani:124102}
and Coupled Cluster theory \cite{PhysRevB.67.045320,heidari:114708,%
  PhysRevB.84.115302}.

Additionally to those approaches, another very promising first-principle
method has recently evolved. This is the Similarity Renormalization Group
(SRG) method, which drives the Hamiltonian to a band- or block-diagonal form
using a continuous series of unitary transformations.  Especially in nuclear
theory, it has successfully been applied to study systems with different
underlying potentials, and it has been used to analyse their binding energy
and other observables \cite{ScottSRG,PhysRevC.75.061001,SRGThreeDim}. Apart
from the free-space approach, where the Hamiltonian is set up with respect to
the zero vacuum state, in recent times another interesting alternative has
been worked out: Making use of the technique of normal-ordering, the SRG
evolution is performed in medium, where 3, \ldots, $A$-body operators can be
approximated using only two-body machinery. This in-medium SRG (IM-SRG) allows
considerable simplifications to the problem and makes numerical calculations
much more efficient \cite{IMSRG}.

 This article is organized as follows: Section \ref{sec:formalism}
 introduces first (\ref{subsec:modelHamiltonian}) the Hamiltonian we
 use to model circular quantum dots, and gives afterwards an overview
 over the SRG (\ref{subsec:SRG}) and DMC (\ref{subsec:DMC})
 formalism. Our results are presented in Section \ref{sec:results}. We
 point out the differences arising from the use of two different
 generators, Wegner's and White's one, and give an overview over the
 ground state energies up to 42 particles. In particular, we compare
 the results obtained with five different \textit{ab initio} methods:
 IM-SRG(2), DMC, Hartree-Fock, Full Configuration Interaction and
 Coupled Cluster. Section \ref{sec:conclusions} concludes our work and
 gives perspectives for future work.

The harmonic oscillator (HO) is ubiquitous in quantum physics. It is
the basis for extremely diverse many-body calculations in nuclear
physics,\cite{HjorthJensen1995} quantum chemistry,\cite{Helgaker2002}
and quantum dot calculations in solid state physics.\cite{Reimann2002}
In these calculations, one typically seeks a few of the lowest
eigenenergies $E_k$ of the system Hamiltonian $H$, and their
corresponding eigenvectors $\psi_k$, viz,
\[ H\psi_k = E_k\psi_k, \quad k = 1,\cdots,k_\text{max}. \] In all
cases, the many-body wave function is expanded in a basis of
eigenfunctions of the HO, and then necessarily truncated to give an
approximation. In fact, the so-called curse of dimensionality implies
that the number of degrees of freedom per particle available is
severely limited. It is clear, that an understanding of the properties
of such expansions, giving \emph{a priori} error estimates on
many-body calculations, is very important. Unfortunately, this is a
neglected topic in the physics literature. In this article, we give a
thorough mathematical and numerical analysis and give practical
convergence estimates of HO basis expansions. It generalizes and
refines the findings of a recent study of one-dimensional systems.\cite{Kvaal2007}

The HO eigenfunctions are popular for several reasons. Many quantum
systems, such as the quantum dot model considered here, are perturbed
harmonic oscillators \emph{per se}, so that the true eigenstates
should be perturbations of the HO states. Moreover, the HO has many
beautiful properties, such as separability, invariance under
orthogonal coordinate changes, and easily computed eigenfunctions, so
that computing matrix elements of relevant operators becomes
relatively simple. The HO eigenfunctions are defined on the
\emph{whole} of $\mathbb{R}^d$ in which the particles live, so that
truncation of the domain is unnecessary. Indeed, this is one of the
main problems with methods such as finite difference or finite element
methods.\cite{RamMohan2002}

@@

We model the quantum dot as a collection of $N$ electrons trapped in an
external harmonic-oscillator potential.\footnote{In this paper, we use atomic
  units where $\hbar$, $m$, $e$, and $4 \pi \epsilon$ are unity, with
  $\epsilon$ denoting the permittivity and $m$ denoting the mass, which may be
  an effective mass.  In vacuum where $m = m_{\text e}$ and
  $\epsilon = \epsilon_0$, energy would be measured in units of hartrees
  $E_{\text h} \equiv m_{\text e} (\alpha c)^2$.  More generally, the units of
  energy are $m_{\text r \vphantom h}^{} E_{\text h}^{} /
  \epsilon_{\text r \vphantom h}^2$ where
  $\epsilon_{\text r} \equiv \epsilon / \epsilon_0$ and
  $m_{\text r} \equiv m / m_{\text e}$.}  Each electron possesses kinetic
energy $t(\bm p) \equiv \bm p^2 / 2$ and experiences an external potential
\begin{align*}
  u(\bm r) \equiv \frac{1}{2} \omega^2 \bm r^2
\end{align*}
where $\omega$ is the angular frequency of the oscillator, $\bm r$ is the
2-dimensional coordinate of an electron, and $\bm p$ is its momentum.  Each
pair of electrons experience a Coulomb repulsion
\begin{align*}
  v(\bm r, \bm r') \equiv \frac{1}{|\bm r - \bm r'|}
\end{align*}
where $\bm r$ and $\bm r'$ are coordinates of a pair of electrons.  Thus, the
full many-body problem is described by the nonrelativistic Hamiltonian
\begin{align*}
  \hat H \equiv \sum_{p} \bigl(
  t(\hat{\bm p}_p) + u(\hat{\bm r}_p)
  \bigr) + \sum_{\{p, q\}} v(\hat{\bm r}_p, \hat{\bm r}_q)
\end{align*}
where $\hat{\bm p}_p$ is the momentum operator and $\hat{\bm r}_p$ is the
position operator for a particle labeled by $p$.  The first summation is
performed over every particle $p$, while the the second summation is performed
over every unordered pair of particles $\{ p, q \}$ (with $p \ne q$).

The full problem is not a trivial one.  We shall first consider a much simpler
problem: a noninteracting Hamiltonian $\hat H^{\text o}$ in which the Coulomb
interaction $v$ is removed:
\begin{align*}
  \hat H^{\text o} \equiv \sum_{p} \bigl(
  t(\hat{\bm p}_p) + u(\hat{\bm r}_p)
  \bigr)
\end{align*}
In this case, the many-body Hamiltonian reduces to $N$ independent
single-particle Hamiltonians, each of the form:
\begin{align*}
  \hat h^{\text o} \equiv
  \frac{\hat{\bm p}^2 + \omega^2 \hat{\bm x}^2}{2}
\end{align*}
This is a 2-dimensional harmonic oscillator problem for a single particle,
whose solutions are well known.  There are, however, various ways to construct
the set of eigenfunctions: we shall use the Fock-Darwin states that conserve
angular momentum about the axis perpendicular to the plane.  The
eigenfunctions have the following analytic form in polar
coordinates:\citep[Appx.\ A]{lohne2010coupled}
\begin{gather*}
  \phi^{\text{fd}}_{n, m_\ell}(r, \varphi)
  \equiv \sqrt\omega
  \phi^{\text{fdr}}_{n, |m_\ell|}(\sqrt \omega r)
  \phi^{\text{fda}}_{m_\ell}(\varphi)
\end{gather*}
where
\begin{align*}
  \phi^{\text{fdr}}_{n, \mu}(\rho) &\equiv
  \sqrt{\frac{2 \cdot n!}{(n + \mu)!}}
  {\mathrm e}^{- \rho^2 / 2} \rho^\mu L_n^{(\mu)}(\rho^2) \\
  \phi^{\text{fda}}_{m_\ell}(\varphi) &\equiv
  \frac{1}{\sqrt{2 \pi}} {\mathrm e}^{i m_\ell \varphi}
\end{align*}
and $L_n^{(\mu)}$ denotes the generalized Laguerre polynomial of degree $n$
with parameter $\mu$.  The states are labelled by two quantum numbers: the
principal quantum number $n \in \{0, 1, \ldots\}$ and the angular momentum
projection $m_\ell \in \mathbb Z$.

Additionally, since electrons are spin-$1/2$ fermions, they can occupy either
of the two spin states $\chi_{+\frac12}$ or $\chi_{-\frac12}$.  Thus, the
single-particle basis states can be constructed as
\begin{align} \label{eq:singleparticlestate}
  \phi_{n, m_\ell, m_{\text s}}(r, \varphi) \equiv
  \phi^{\text{fd}}_{n, m_\ell}(r, \varphi) \chi_{m_{\text s}}
\end{align}
with the spin projection $m_{\text s} \in \bigl\{-\frac12, +\frac12\bigr\}$.

\begin{figure}
\includegraphics[width=.48\textwidth]{fig-shell-structure}
\caption{Structure of the single-particle states in a noninteracting harmonic
  oscillator.}
\label{fig:shell-structure}
\end{figure}
The eigenvalues of the single-particle Hamiltonian $\hat h^{\text o}$ are
given by
\begin{align*}
  \varepsilon_{n, m_\ell, m_{\text s}} \equiv (2 n + |m_\ell| + 1) \omega
\end{align*}
Note that they are degenerate with respect to the spin projection
$m_{\text s}$ since our Hamiltonian $\hat h^{\text o}$ does not distinguish
between them.  From this equation, we can see that the single-particle basis
states form a shell structure as shown in Figure \ref{fig:shell-structure}.
Every state within a given shell is degenerate and states within a given shell
satisfy $2 n + |m_\ell| = k$ for some nonnegative integer $k$, which we call
the \textit{shell index}.  The shells are equidistant from each other, with a
spacing equal to the frequency $\omega$.

The many-body Hamiltonian $\hat H^{\text o}$ is the sum of the single-particle
Hamiltonians $\hat h^{\text o}_p$ over all particles $p$, where $p$ may now be
treated as a tuple of all three single-particle quantum numbers:
$(n, m_\ell, m_{\text s})$.  From the states $\phi$ in
\eqref{eq:singleparticlestate} we can construct the eigenstates of the
many-body Hamiltonian $\hat H^{\text o}$ in the form of Slater determinants
that automatically satisfy the Pauli exclusion principle:
\begin{align*}
  \Phi_{p_1, \ldots, p_N}(\bm r_1, \ldots, \bm r_N) \equiv
  \frac{1}{\sqrt{N}} \left|
  \begin{matrix}
    \phi_{p_1}(\bm r_1) & \cdots & \phi_{p_N}(\bm r_1) \\
    \vdots & \ddots & \vdots \\
    \phi_{p_1}(\bm r_N) & \cdots & \phi_{p_N}(\bm r_N) \\
  \end{matrix}
  \right|
\end{align*}
The energy of these states is given by the sum of the single-particle
energies:
\begin{align*}
  E^{\text o}_{p_1, \ldots, p_N} \equiv
  \varepsilon_{p_1} + \cdots + \varepsilon_{p_N}
\end{align*}
For ground states, we may have degeneracies due to the shell structure of the
single-particle basis.  However, if $N = k_{\text f} (k_{\text f} + 1)$ for
some nonnegative integer $k_{\text f}$, there are just enough particles to
form a closed-shell system.  In this case, there is a unique, well-isolated
ground state.  The values of $N$ for which this occurs are typically referred
to as \textit{magic numbers} and we call $k_{\text f}$ the \textit{number of
  filled shells}.  This completes the solution of the noninteracting problem.

The difficulty in solving the full Hamiltonian $\hat H$ stems from the fact
that the exact solution cannot be expressed in terms of a single Slater
determinant due to the presence of interactions.  Hence, it is not possible to
reduce the problem to that of a single particle as we had done previously.

@@

Since Slater determinants are rather cumbersome to write out in full, we shall
instead work in the more convenient formalism of second quantization:
\begin{align*}
  |\Phi_{p_1, \ldots, p_N}\rangle =
  \hat a_{p_1}^\dagger \cdots \hat a_{p_N}^\dagger |\varnothing\rangle
\end{align*}
Here, $\hat a_p^\dagger$ is the creation operator for the single-particle
state $p$ and $|\varnothing\rangle$ is the (true) vacuum state that is devoid
of particles.  Analogously, $\hat a_p$ is the annihilation operator for the
state $p$.

Operators can be rewritten in this formalism as well.  The one-body kinetic
energy operator, defined by
\begin{align*}
  \hat T \equiv \sum_p \hat t(\hat{\bm p}_p)
\end{align*}
can be rewritten in second quantization as
\begin{align*}
  \hat T = \sum_{p, q} T_{p, q}^{} \hat a_p^\dagger \hat a_q^{}
\end{align*}
where the operator matrix element
$T_{p, q} \equiv \langle \Phi_p | \hat T | \Phi_q \rangle$.  This can likewise
be extended for both $\hat U$ and $\hat V$,
\begin{align*}
  \hat U &= \sum_{p, q} U_{p, q}^{} \hat a_p^\dagger \hat a_q^{} \\
  \hat V &= \frac{1}{4} \sum_{p, q, r, s} V_{p, q, r, s}^{}
  \hat a_p^\dagger \hat a_q^\dagger \hat a_s^{} \hat a_r^{}
\end{align*}
where $U_{p, q} \equiv \langle \Phi_p | \hat U | \Phi_q \rangle$ and
$V_{p, q, r, s} \equiv \langle \Phi_{p, q} | \hat V | \Phi_{r, s} \rangle$.
Note that due to the antisymmetry present in the Slater determinants, we have
the property that exchanging either $p$ with $q$ or $r$ with $s$ causes
$V_{p, q, r, s}$ to reverse in sign.

>>>>

\documentclass[amsmath, amssymb, aps, floatfix, nofootinbib, preprintnumbers,showpacs, superscriptaddress, twocolumn]{revtex4-1}
\usepackage{bm}
\usepackage[cmyk]{xcolor}
\usepackage{dcolumn}
\usepackage{epsfig}
\usepackage{float}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{subfigure}

\newcommand{\D}{\operatorname{d\!}}
\newcommand{\ket}[1]{\left| #1 \right>}
\newcommand{\bra}[1]{\left< #1 \right|}
\newcommand{\braket}[2]{\left\langle #1 | #2\right\rangle}
\newcommand{\normord}[1]{\mathopen: #1 \mathclose:}

@@

\begin{abstract}
  We present and compare several many-body methods used to study a quantum dot
  system.  We calculate the approximate ground state energy of electrons in a
  circularly symmetric potential well using a harmonic oscillator basis
  optimized via Hartree-Fock (HF).  We further improve the ground state energy
  using in-medium similarity renormalization group (IM-SRG).  Additionally, we
  calculate the addition and removal energies using quasidegenerate
  perturbation theory (QDPT).  Our results are benchmarked against quasi-exact
  diffusion Monte Carlo (DMC) and coupled cluster (CC) calculations.
\end{abstract}

@@

Understanding the behavior of strongly confined electrons is of fundamental
interest for solving many-body problems.  Quantum dots, e.g. electrons
confined in semiconducting heterostructures, are of particular interest since
they exhibit, due to their small size, discrete quantum levels.  Under these
conditions, typical quantum phenomena like tunneling, entanglement and
magnetization can all be observed \cite{reimann2002,engel1993}.  Since quantum
dots are manufactured and designed artificially at the laboratory, their
quantum levels can be arbitrarily tuned by changing the external field or the
size and shape of the system.  As a consequence, quantum dots provide a high
level of control for the dynamics and correlation of the electrons, which
makes them perfectly suited to study quantum effects in practice.  Since their
ground state shows similar shell structures and magic numbers as seen for
atoms and nuclei \cite{tarucha1996}, these systems give the opportunity to
study electronic systems without the presence of a nucleus affecting the
electrons.  Apart from their relevance for theoretical research in quantum
physics, quantum dots offer a wide variety of applications: In particular,
their electrical and optical properties make them attractive for the use in
laser technology \cite{strauf2010,5075760} and solar cells
\cite{jenks:013111,doi:10.1021/cr900289f}, but they are also used in quantum
computers\cite{PhysRevA.57.120} and medical imaging \cite{Ben-Ari02042003}.

In order to properly understand the properties of quantum dots and make
theoretical predictions of their behavior in various applications, it is
necessary to study features like ground state energy and correlation
effects. Since there generally does not exist an analytical solution for
quantum dots except for ones containing only two electrons or with specific
values of the external field \cite{PhysRevA.48.3561}, the development of
appropriate numerical few- and many-body methods is required to solve the
problem.

Several \textit{ab initio} methods have been applied to these systems, in
particular variational and diffusion Monte Carlo (VMC and DMC)
\cite{PhysRevB.68.035304,%
  PhysRevB.62.8120,PhysRevB.84.115302,PhysRevB.54.4780}, large-scale
diagonalization/full configuration interaction (FCI)
\cite{JJAP.36.3924,PhysRevB.56.6428,Kvaalcode,rontani:124102} and coupled
cluster (CC) theory \cite{PhysRevB.67.045320,heidari:114708,%
  PhysRevB.84.115302}.

Additionally to those approaches, another very promising first-principles
method has recently been developed.  The similarity renormalization group
(SRG) is a family of methods that drives the Hamiltonian toward a band- or
block-diagonal form through a continuous sequence of unitary transformations,
reducing the coupling between a subspace of interest -- such as the ground
state -- and the remaining Hilbert space.  It has successfully been applied to
study systems with different underlying potentials and to analyse their
binding energy and other observables, especially in nuclear theory
\cite{ScottSRG,PhysRevC.75.061001,SRGThreeDim}.  Aside from usual
\emph{free-space} SRG, where the Hamiltonian is constructed with respect to
the vacuum state with no physical particles, it is possible to evolve a
Hamiltonian via SRG \emph{in medium} through the normal-ordering technique,
which enables the Hamiltonian to be constructed with respect to some reference
state containing a finite number of particles.  In-medium SRG (IM-SRG) can
further take advantage of the normal-ordering process to account for 3-body
forces in an approximate manner without resorting to the full, computationally
intensive 3-body machinery.  The method allows considerable simplifications to
the problem and makes numerical calculations much more efficient \cite{IMSRG}.

This article is organized as follows: Section \ref{sec:formalism} introduces
first (\ref{subsec:modelHamiltonian}) the Hamiltonian we use to model circular
quantum dots, and gives afterwards an overview of the IM-SRG
(\ref{subsec:SRG}) and DMC (\ref{subsec:DMC}) formalisms.  Our results are
presented in Section \ref{sec:results}.  We point out the differences arising
from the use of two different generators (Wegner and White generator), and
give an overview of the ground state energies up to 42 particles. In
particular, we compare the results obtained with five different \textit{ab
  initio} methods: IM-SRG(2), DMC, Hartree-Fock, FCI and CC. Section
\ref{sec:conclusions} concludes our work and gives perspectives for future
work.

The harmonic oscillator (HO) is a ubiquitous system in quantum physics. It is
the basis for an extremely diverse set of many-body calculations -- in nuclear
physics \cite{HjorthJensen1995}, quantum chemistry \cite{Helgaker2002}, and
quantum dot calculations of solid state physics \cite{Reimann2002}. In these
calculations, one typically seeks a few of the lowest eigenvalues $E_k$ and
eigenvectors of a many-body Hamiltonian $\hat{H}$ that describes the full
system by solving the Schr\"odinger equation,
\begin{equation*}
\hat{H} \ket{\psi_k} = E_k \ket{\psi_k}, \quad k \in \{1, 2, \ldots, k_{\text{max}}\}.
\end{equation*}
In all cases, the many-body wave function is expanded in a basis of
eigenstates of the one-body HO system.  Necessarily, in order to perform
numerical calculations the basis must be truncated to give an approximation.
In particular, the so-called ``curse of dimensionality'' implies that for a
system with many particles the number of available degrees of freedom per
particle is severely limited by the finiteness of a computer.

[[This paragraph seems to be talking about a different paper]]
It is clear, that an understanding of the properties of such expansions,
giving \emph{a priori} error estimates on many-body calculations, is very
important. Unfortunately, this is a neglected topic in the physics
literature. In this article, we give a thorough mathematical and numerical
analysis and give practical convergence estimates of HO basis expansions. It
generalizes and refines the findings of a recent study of one-dimensional
systems.\cite{Kvaal2007}

The HO eigenfunctions are popular for several reasons. Many quantum systems,
such as the quantum dot model considered here, can be considered perturbed
harmonic oscillators \textit{per se}, thus the true eigenstates should be
perturbations of the HO eigenstates.  Moreover, the HO has many beautiful
properties, such as separability, invariance under orthogonal coordinate
changes, and easily computable eigenfunctions, so that computing matrix
elements of relevant operators becomes relatively simple.  Since HO
eigenfunctions are defined on the \emph{entire} space in which particles live,
a truncation of the spatial domain is unnecessary.  This avoids one of the
main problems with approaches such as finite difference or finite element
methods.\cite{RamMohan2002}

@@

We model the quantum dot as a collection of $N$ electrons trapped in an
external harmonic-oscillator potential.\footnote{In this paper, we use atomic
  units where $\hbar$, $m$, $e$, and $4 \pi \epsilon$ are set to 1.  Here,
  $\epsilon$ denotes the permittivity and $m$ denotes the mass, which could be
  an effective mass rather than the true electron mass.  In effect, the unit
  of energy in this article is
  $E_{\text h} (m / m_{\text e}) / (\epsilon / \epsilon_0)^2$ where
  $E_{\text h} \equiv m_{\text e} (\alpha c)^2$, the hartree unit of energy.}
Each electron possesses kinetic energy $t(\bm p) \equiv \bm p^2 / 2$ and
experiences an external potential $u(\bm r) \equiv \omega^2 \bm r^2 / 2$,
where $\omega$ is the angular frequency of the oscillator, $\bm r$ is the
2-dimensional position vector of an electron, and $\bm p$ is its linear
momentum.  Each pair of electrons experience a Coulomb repulsion
$v(\bm r, \bm r') \equiv |\bm r - \bm r'|^{-1}$ where $\bm r$ and $\bm r'$ are
their positions.  The full many-body problem is described by the
nonrelativistic Hamiltonian
\begin{align*}
  \hat H \equiv \sum_{p} \bigl(
  t(\hat{\bm p}_p) + u(\hat{\bm r}_p)
  \bigr) + \sum_{p < q} v(\hat{\bm r}_p, \hat{\bm r}_q)
\end{align*}
where $\hat{\bm p}_p$ and $\hat{\bm r}_p$ respectively denote the momentum and
position operators of a particle $p$.  The first summation is performed over
every particle, while the the second summation is performed over every
unordered pair of distinct particles (the ordering of $p$ is arbitrary).

The full problem is not a trivial one.  We shall first consider a much simpler
problem: a noninteracting Hamiltonian $\hat H^{\text o}$ in which the Coulomb
interaction $v$ is removed:
\begin{align} \label{eq:onebodyhamiltonian}
  \hat H^{\text o} \equiv \sum_p \hat h^{\text o}_p
  \equiv \sum_{p} \bigl(
  t(\hat{\bm p}_p) + u(\hat{\bm r}_p)
  \bigr)
\end{align}
In this case, the many-body Hamiltonian reduces to $N$ independent
single-particle Hamiltonians, each of the form:
\begin{align*}
  \hat h^{\text o} \equiv
  \frac{\hat{\bm p}^2 + \omega^2 \hat{\bm x}^2}{2}
\end{align*}
This is a 2-dimensional harmonic oscillator problem for a single particle,
whose solutions are well known.  There are various ways to construct the set
of eigenfunctions: we shall use the Fock-Darwin states that conserve angular
momentum $\hat{L}_{\text{z}}$ about the $z$-axis (perpendicular to the 2D
plane).  The eigenfunctions have the following analytic form in polar
coordinates:\citep[Appx.\ A]{lohne2010coupled}
\begin{gather*}
  \phi^{\text{FD}}_{n, m_\ell}(r, \varphi)
  \equiv \sqrt\omega
  \phi^{\text{FD.R}}_{n, |m_\ell|}(\sqrt \omega r)
  \phi^{\text{FD.A}}_{m_\ell}(\varphi)
\end{gather*}
where
\begin{align*}
  \phi^{\text{FD.R}}_{n, \mu}(\rho) &\equiv
  \sqrt{\frac{2 \cdot n!}{(n + \mu)!}}
  {\mathrm e}^{- \rho^2 / 2} \rho^\mu L_n^{(\mu)}(\rho^2) \\
  \phi^{\text{FD.A}}_{m_\ell}(\varphi) &\equiv
  \frac{1}{\sqrt{2 \pi}} {\mathrm e}^{i m_\ell \varphi}
\end{align*}
and $L_n^{(\mu)}$ denotes the generalized Laguerre polynomial of degree $n$
with parameter $\mu$.  The states are labeled by two quantum numbers: the
principal quantum number $n \in \{0, 1, \ldots\}$ and the angular momentum
projection $m_\ell \in \mathbb Z$.

Additionally, since electrons are spin-$1/2$ fermions, they can occupy either
of the two spin states $\chi_{-\frac12}$ or $\chi_{+\frac12}$.  Thus, the
single-particle basis states can be constructed as
\begin{align} \label{eq:singleparticlestate}
  \phi_{n, m_\ell, m_{\text s}}(r, \varphi) \equiv
  \phi^{\text{FD}}_{n, m_\ell}(r, \varphi) \chi_{m_{\text s}}
\end{align}
with the spin projection $m_{\text s} \in \bigl\{-\frac12, +\frac12\bigr\}$.

\begin{figure}
\includegraphics[width=.48\textwidth]{fig-shell-structure}
\caption{Structure of the lowest 12 single-particle states in a noninteracting
  harmonic oscillator of frequency $\omega = 1$.  Each box represents a
  single-particle state, and the arrows denote the spin of the states.}
\label{fig:shell-structure}
\end{figure}
The eigenvalues of the single-particle Hamiltonian $\hat h^{\text o}$ are
given by
\begin{align*}
  \varepsilon_{n, m_\ell, m_{\text s}} \equiv (2 n + |m_\ell| + 1) \omega
\end{align*}
which satisfies
$\hat{h}^{\text o} \ket{\phi_{n, m_\ell, m_{\text s}}} = \varepsilon_{n,
  m_\ell, m_{\text s}} \ket{\phi_{n, m_\ell, m_{\text s}}}$.  Note that the
states are degenerate with respect to the spin projection $m_{\text s}$ since
our Hamiltonian $\hat h^{\text o}$ does not distinguish between them.  From
this equation, we can deduce the shell structure as shown in Figure
\ref{fig:shell-structure}.  It is notable that there are degeneracies beyond
that of the spin pairs: each shell contains a set of states that satisfy the
equation $2 n + |m_\ell| = k$ for some nonnegative integer $k$, which we
conveniently call the \textit{shell index}.  The shells are equidistant from
each other, with a spacing equal to the frequency $\omega$.

The many-body Hamiltonian $\hat H^{\text o}$ is the sum of the single-particle
Hamiltonians $\hat h^{\text o}_p$ over all particles $p$, where $p$ may now be
treated as a tuple of all three single-particle quantum numbers:
$(n, m_\ell, m_{\text s})$.  From the states $\phi$ in
\eqref{eq:singleparticlestate} we can construct the an eigenstate of the
many-body Hamiltonian $\hat H^{\text o}$ in the form of a Slater determinant
$\varnothing_{p_1, \ldots, p_N}$:
\begin{align*}
  \varnothing_{p_1, \ldots, p_N}(\bm r_1, \ldots, \bm r_N) \equiv
  \frac{1}{\sqrt{N}} \left|
  \begin{matrix}
    \phi_{p_1}(\bm r_1) & \cdots & \phi_{p_N}(\bm r_1) \\
    \vdots & \ddots & \vdots \\
    \phi_{p_1}(\bm r_N) & \cdots & \phi_{p_N}(\bm r_N) \\
  \end{matrix}
  \right|
\end{align*}
This automatically satisfies the Pauli exclusion principle, ensuring that the
Fermi statistics are satisfied.  The energy of these states is given by the
sum of the single-particle energies:
\begin{align*}
  E^{\text o}_{p_1, \ldots, p_N} =
  \varepsilon_{p_1} + \cdots + \varepsilon_{p_N},
\end{align*}
which satisfies
$\hat{H}^{\text o} \ket{\varnothing_{p_1, \ldots, p_N}} = E^{\text o}_{p_1, \ldots,
  p_N} \ket{\varnothing_{p_1, \ldots, p_N}}$.  For ground states, we may have
degeneracies due to the shell structure of the single-particle basis.
However, if $N = k_{\text F} (k_{\text F} + 1)$ for some nonnegative integer
$k_{\text F}$, there are just enough particles to form a closed-shell system.
In this case, there is a unique, well-isolated ground state.  The values of
$N$ for which this occurs are typically referred to as \textit{magic numbers}
and we call $k_{\text F}$ the \textit{number of filled shells} (or ``Fermi
level'').  This completes the solution of the noninteracting problem.

The difficulty in solving the full Hamiltonian $\hat H$ stems from the fact
that the exact solution cannot be expressed in terms of a single Slater
determinant due to the presence of interactions.  Hence, it is not possible to
trivially reduce the problem to that of a single particle as we had done
previously.

@@

Since Slater determinants are rather cumbersome to write out in full, we shall
instead work in the more convenient formalism of \textit{second quantization}:
\begin{align*}
  \ket{\varnothing_{p_1, \ldots, p_N}} =
  \hat a_{p_1}^\dagger \cdots \hat a_{p_N}^\dagger \ket{\varnothing}
\end{align*}
Here, $\varnothing$ is the vacuum state that is devoid of particles and
$\hat a_p^\dagger$ is a \textit{creation operator} for the single-particle
state $p$.  If the state $p$ is not yet occupied by a particle,
$\hat a_p^\dagger$ creates such a particle.  Otherwise, it collapses the
entire many-body state to $0$.  The Hermitian conjugate of $\hat a_p^\dagger$
is $\hat a_p$, which is an \textit{annihilation operator} for the state $p$
that has the opposite effect.  For fermions, the field operators $\hat a_p$
and $\hat a_p^\dagger$ satisfy the essential anticommutation relations,
\begin{align*}
  &\{\hat a_p, \hat a_q\} = \{\hat a_p^\dagger, \hat a_q^\dagger\} = 0, &
  &\{\hat a_p, \hat a_q^\dagger\} = \delta_{p, q}.
\end{align*}
Operators can be rewritten in this formalism as well.  The one-body
Hamiltonian $H^{\text o}$ defined by \eqref{eq:onebodyhamiltonian} can be
rewritten in second quantization as
\begin{align} \label{eq:second_quantized_one_body_hamiltonian}
  \hat H^{\text o} = \sum_{p, q} H_{p, q}^{\text o} \hat a_p^\dagger \hat a_q^{}
\end{align}
where the \textit{matrix element}
$H_{p, q}^{\text o} \equiv \bra{\varnothing_p} \hat H^{\text o} \ket{\varnothing_q}$.  This
can likewise be extended to $\hat V$,
\begin{align} \label{eq:second_quantized_two_body_hamiltonian}
  \hat V = \frac{1}{4} \sum_{p, q, r, s} V_{p, q, r, s}^{}
  \hat a_p^\dagger \hat a_q^\dagger \hat a_s^{} \hat a_r^{}
\end{align}
where $U_{p, q} \equiv \langle \varnothing_p | \hat U | \varnothing_q \rangle$ and
$V_{p, q, r, s} \equiv \langle \varnothing_{p, q} | \hat V | \varnothing_{r, s} \rangle$.
Note that due to the antisymmetry present in the Slater determinants, we have
the property that exchanging either $p$ with $q$ or $r$ with $s$ causes
$V_{p, q, r, s}$ to reverse in sign, leading to a $4$-fold symmetry that is
canceled out by the factor of $1/4$.  For this reason, the two-body matrix
element $V_{p, q, r, s}$ is commonly referred to as the \emph{antisymmetrized}
matrix element.  This is not to be confused with the non-antisymmetric
interaction integral defined by:
\begin{align*}
  \tilde V_{p, q, r, s} \equiv \int \phi_p(\bm r) \phi_q(\bm r') v(\bm r, \bm r') \phi_r(\bm r) \phi_s(\bm r') \D \bm r \D \bm r'
\end{align*}
The antisymmetrized matrix element is computed from the interaction integral
by $V_{p, q, r, s} = \tilde V_{p, q, r, s} - \tilde V_{p, q, s, r}$.  In our
work, we compute the interaction numerically via the OpenFCI software, which
numerical integrates the Coulomb interaction in the center-of-mass frame
[[cite Kvaal's paper; also, is the description correct?]].
